### 2025-06-02 Manuscript/Lower bounds of the success probability in quantum state exclusion for general ensembles.tex
Stuff

\documentclass[12pt,letterpaper]{article}
\usepackage{fullpage}
\usepackage[top=1.5cm, bottom=3.5cm, left=2.2cm, right=2.2cm]{geometry}
\usepackage{amsmath,amsthm,amsfonts,amssymb,amscd, esint}
\usepackage{lastpage}
\usepackage{enumerate}
\usepackage{fancyhdr}
\usepackage{mathrsfs}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{hyperref}
\usepackage[english]{babel}
\usepackage{lipsum}
\usepackage[table,xcdraw]{xcolor}
\usepackage{enumitem}
\usepackage{overpic}
\usepackage{float}
\usepackage{yfonts}
\usepackage{braket}
\usepackage{dsfont}
\usepackage{tikz}
\usepackage{wrapfig}
\usepackage{url}
\usepackage{natbib}
\usepackage[normalem]{ulem}
\usepackage{multicol}
\useunder{\uline}{\ul}{}

@@

\thispagestyle{empty}
\mbox{} 
\newpage
\thispagestyle{empty}
\vspace*{\fill} % Push content to vertical center
\begin{flushright}
    \emph{Ab ovo usque ad mala.}\\[1em]
    \textbf{Horace}
\end{flushright}
\vspace*{\fill} 
\newpage
\thispagestyle{empty}
\mbox{} 
\newpage
\thispagestyle{empty}
\cfoot{}
\tableofcontents

\newpage
\pagestyle{fancyplain}
\headheight 35pt
\lhead{\NetIDb}    
\rhead{\subject}
\cfoot{\small\thepage}
\headsep 1.5em
\setcounter{page}{1}
\begin{center}
\begin{abstract}
Given a quantum state known to be prepared from an ensemble of two or more states, quantum state exclusion aims to rule out the possibility that it was prepared in a particular state from the ensemble. Using the known solution for group generated ensembles \cite{MainPaper}, we study this result as a lower bound for randomly generated ensembles via semidefinite programming.
\end{abstract}
\end{center}
%Keywords: \emph{SDP, Quantum state exclusion , \textcolor{blue}{Add keywords}}.
\section{Introduction}

@@

In the quantum realm, a similar line of reasoning can be applied in some problmes. One of the fundamental tasks in quantum information theory is the identification of an unknown quantum state known to belong to a given ensemble of possible states. This task is referred to as Quantum State Discrimination (\gls{qsd}), and it has been extensively studied in the literature \cite{DiscriminationArticle}. However, instead of aiming to identify the exact state, we may take another approach by applying the hypothesis exclusion idea. That is, given a quantum state known to be one among a certain ensemble, the goal becomes to conclusively rule out the possibility of the unknown state being a specific member of that ensemble. This alternative task is known as Quantum State Exclusion (\gls{qse}) \cite{PhysicsExclusionSource}, and it offers a distinct framework with its own set of advantages \cite{piani2020quantum} with respect to its counterpart \gls{qsd}.

We can approach \gls{qse} task under several protocols, depending on the context within we perform the hypothesis exclusion we can prefer one or another. In this project, we focus on two main protocols, Minimum Error (\gls{me}) and Zero Error (\gls{ze})\footnote{Also known as \emph{unambiguous exclusion}.}, i.e. the protocols for which analytical results have been presented \cite{MainPaper}. In the Minimum Error scenario, the goal is to minimize the probability of mistakenly excluding the actual prepared state. In contrast, the Zero Error approach seeks to exclude a state with absolute certainty, although this procedure sometimes yields an inconclusive result when perfect exclusion cannot be accomplished.

While \gls{qsd} has seen significant progress over the years \cite{bae2015quantum, helstromBook}, \gls{qse} has remained relatively underexplored until recently. New analytical results have emerged \cite{MainPaper} for both the \gls{me} and \gls{ze} protocols, particularly when the ensemble of quantum states possesses a certain degree of symmetry. In particular, when the ensemble is generated through the action of a finite algebraic group, the problem becomes more tractable, highlighting \gls{qse} as a worthwhile object of study.

Although the tasks of exclusion and discrimination coincide for ensembles containing only two states\footnote{Since for the two states case excluding one necessarily implies identifying the other.}, when dealing with ensembles of three or more states, the two problems diverge in both approach and complexity. Though this work primarily focuses on \gls{qse}, we include a comparative discussion of both tasks and protocols to further motivate the interest in \gls{qse} and discuss the main differences with \gls{qsd} in Appendix \ref{appendixComparisonQSDvsQSE}. One of the key features of \gls{qse} is the possibility of achieving \emph{perfect exclusion}, i.e. confidently ruling out a candidate state with complete certainty that it is not the unknown state. In contrast, \emph{perfect discrimination} is generally not achievable, meaning that one cannot identify the unknown state with absolute certainty in most settings\cite{OptimalitySRM}.

@@

Building upon the recent results for the probability of a correct exclusion for group-generated ensembles \cite{MainPaper},\footnote{A proper definition of a \emph{group generated ensemble} will be given in the \emph{Group generated ensembles} section \ref{sectionGroupGeneratedEnsemble}, for the time being, we can identify them as ensembles that comprehend set of states that exhibit a certain symmetry.} this project undertakes a numerical investigation into \gls{qse}, highlighting the analytical results from group-generated ensembles as lower bounds for more general cases. The project comences with a preliminaries section that outlines the mathematical framework underlying the \gls{qse} task, along with the essential tools used in our analysis. We then present results for the probability of correct hypothesis exclusion in ensembles consisting of three quantum states, the simplest case. Furthermore, we explore the distribution of correct exclusion probabilities for randomly generated ensembles sampled according to Haar measure.\footnote{The importance of the generic ensembles generation being performed according to Haar measure is a key for the numerical analysis and it is discussed in the \emph{Methodology} section (\ref{sectionMethodology})}

To carry out our numerical analysis, we employ \gls{sdp} techniques, which allow us to cast the \gls{qse} task as a convex optimization problem. The \gls{sdp} approach has proven particularly effective for problems of this kind \cite{bandyopadhyay2014conclusive}. Finally, ve conclude by providing some theoretical intuition and insights based on our results.

@@

Unlike projective measurements, the operators $\Pi_i$ in a POVM are not necessarily orthogonal projectors. This allows the number of measurement outcomes to exceed the dimensionality of the Hilbert space, extending the previos orthogonal measurement to a more generic status and making POVMs particularly useful in tasks such as \gls{qse} and \gls{qsd}.

@@

\hspace{20pt}Let $\left\{(\rho_i, \eta_i)\right\}_{i=1}^n$ be an ensemble of $n$ quantum states, where each $\rho_i$ denotes a pure state density matrix\footnote{This formulation hols true for mixed states but the project will only discuss the pure state scenario.}, i.e., $\rho_i = \ket{\psi_i}\bra{\psi_i}$, and $\eta_i$ represents the prior probability of occurrence of the state $\rho_i$. Let $\tilde{\rho}$ be a state from this ensemble. Our objective is to develop a procedure to identify another state $\rho_j \in \left\{\rho_i\right\}_{i=1}^n$, such that $\rho_j \neq \tilde{\rho}$.

Quantum measurements are described by a set of \glspl{povm}, denoted by $\left\{\Pi_i\right\}_{i=1}^n$, acting on the Hilbert space $\mathcal{H}$ of the quantum system. Here we present the two studied protocols for \gls{qse}: minimum-error (\gls{me}) and zero-error (\gls{ze}).

As previously discussed the goal of the \gls{me} protocol is to minimize the probability of incorrectly excluding the unknown state from our hypothesis. If we formulate it as an \gls{sdp}, the \gls{sdp} minimization of the error probability reads,\footnote{Note that the \gls{sdp} formulations of quantum state discrimination may differ from the exclusion ones by interchanging minimization and maximization problems.}
\begin{align*}
	P_{\text{\gls{me}}}^e = \min_{\left\{\Pi_i\right\}} &\sum_{i=1}^n \eta_i \tr(\Pi_i \rho_i),\\
	\text{subject to} \quad & \sum_{i=1}^n \Pi_i = \mathds{1}, \quad \Pi_i \geq 0 \quad \forall i \in \{1, \dots, n\}.
\end{align*}

@@

In the case of the \gls{ze} protocol, the objective is to exclude with total accuracy a state known to be different from the unkown state. However, the protocol might lead to inconclusive result. In order to do so, the \gls{povm}s must also satisfy an unambiguity condition, i.e. each measurement operator $\Pi_i$ must be orthogonal to the corresponding state $\rho_i$. In other words,
\begin{align*}
\tr(\Pi_i \rho_i) = 0 \quad \forall i \in \{1, \dots, n\}.
\end{align*}

@@

If the measurement yields the outcome $\Pi_?$, i.e., the "?" label, the result is inconclusive.

@@

This reformulation reveals that if two ensembles $A$ and $B$ share the same Gram matrix, then their exclusion problems under both the \gls{me} and \gls{ze} protocols are equivalent. That is, their optimal success and error probabilities coincide. Consequently, we focus on the Gram matrix to study exclusion problems rather than the explicit state representations.

@@

\hspace{20pt}As we have mentioned before, analytical results have been found for the exclusion task in both protocols \gls{me} and \gls{ze} for ensembles that exhibit a certain degree of symmetry\cite{MainPaper}, that is for \emph{group-generated} ensembles. In this section we present a proper definition for a group-generated ensemble as well as a brief discussion on this ensembles generated by a cyclic group, a key type of grop-generated ensembles in our analysis.

@@

forms a group generated ensemble based on the cyclic group $\mathbb{Z}/n\mathbb{Z}$. This type of ensemble is of particular interest in our study and will be explored in more detail at the end of this section. Additionally, in Appendix \ref{appendixComputationGroupGeneratedGramMatrices} we present a general procedure to compute the Gram Matrix for any group-generated ensemble.

\subsubsection*{The $\mathbb{Z}/n\mathbb{Z}$ group-generated ensemble}

\hspace{20pt}Here we discuss the particular case for a $\mathbb{Z}/n\mathbb{Z}$ group-generated ensemble a key type of ensemble for our project.\footnote{The relevance of this type of ensembles relies on the fact that for a prime number of states, the possible group-generated group ensemble } Let $\mathcal{U} \in U(n)$ be an $n \times n$ unitary matrix satisfying $\mathcal{U}^n = \mathds{1}$, and let $\ket{\psi}$ be the seed state. The Gram matrix $\mathcal{G}^{\mathbb{Z}/n\mathbb{Z}}$ elements associated with the ensemble $\left\{\mathcal{U}^i\ket{\psi}\right\}_{i=0}^{n-1}$ are,
\begin{align*}
	\mathcal{G}^{\mathbb{Z}/n\mathbb{Z}}_{i,j} = \sqrt{\eta_i \eta_j}\braket{\mathcal{U}^i\psi | \mathcal{U}^j\psi} = \sqrt{\eta_i \eta_j}\braket{\psi | \mathcal{U}^{j-i} | \psi} = \sqrt{\eta_i \eta_j}\braket{\mathcal{U}^{j-i}}_{\psi},
\end{align*}

@@

This shows the circulant structure of $\mathcal{G}$ for a equal prior probability case, where each row is a cyclic permutation of the previous one. In other words, the Gram matrices of $\mathbb{Z}/n\mathbb{Z}$-generated ensembles are circulant matrices when taking equal priors probabilities. This structure is particularly useful, as circulant matrices can be diagonalized by the discrete Fourier basis, greatly simplifying many computations~\cite{circulantMatrices}.

@@

Finally, we denote the cardinality of a finite group $G$, i.e. the number of its elements of the group, as $|G|$. In scenarios with uniform prior probabilities, this implies,
\begin{align*}
	\eta_i = \frac{1}{|G|}\quad \forall i\in\{1,...,n\}.
\end{align*}
\subsection{Previous Results}\label{sectionPreviousResults}

\hspace{20pt}Significant progress has been made in the study of the state exclusion task for group-generated ensembles. In particular, the work \emph{Quantum State Exclusion for Group-Generated Ensembles of Pure States}\cite{MainPaper} presents exact analytical expressions for the success and error probabilities under both the Minimum Error (\gls{me}) and Zero Error (\gls{ze}) exclusion protocols. In this section we present this results and comment some key aspects of this results regarding our project, in particular we present the concept of \emph{perfect exclusion region} and we discuss the case of ensembles with a prime number of states.

Let $\mathcal{G}^G$ denote the Gram matrix of a group-generated ensemble $\left\{ \left( \rho_i, \frac{1}{|G|} \right) \right\}_{i=1}^{|G|}$, where $G$ is a finite group of size $|G|$, and let $\{\lambda_i\}_{i=1}^{|G|}$ be the set of eigenvalues of $\mathcal{G}^G$, the Gram Matrix of the group. Assume the eigenvalues $\left\{\frac{\lambda_i}{|G|}|1\leq i \leq n\right\}$ are ordered increasingly,
\begin{align*}
	i \leq j \quad \Leftrightarrow \quad \lambda_i \leq \lambda_j, \quad \forall i,j \in \{1, \dots, |G|\},
\end{align*}

@@

The exclusion probabilities under the Minimum Error protocol are given by,
\begin{align*}
	P_{\gls{me}}^e &= \max\left\{ 0, \left( \frac{ \sqrt{\lambda_{|G|}} - \sum_{i<|G|} \sqrt{\lambda_i} }{|G|} \right)^2 \right\}, \\
	P_{\gls{me}}^s &= \min\left\{ 1, 1 - \left( \frac{ \sqrt{\lambda_{|G|}} - \sum_{i<|G|} \sqrt{\lambda_i} }{|G|} \right)^2 \right\}.
\end{align*}

For the Zero Error protocol, the corresponding expressions arem
\begin{align*}
	P_{\gls{ze}}^e &= \max\left\{ 0, \frac{ \sum_{i=1}^{|G|} \sqrt{\lambda_i} \left( \sqrt{\lambda_{|G|}} - \sum_{j<|G|} \sqrt{\lambda_j} \right) }{|G|} \right\}, \\
	P_{\gls{ze}}^s &= \min\left\{ 1, 1 - \frac{ \sum_{i=1}^{|G|} \sqrt{\lambda_i} \left( \sqrt{\lambda_{|G|}} - \sum_{j<|G|} \sqrt{\lambda_j} \right) }{|G|} \right\}.
\end{align*}

A key feature of these results is their independence from the specific group generating the ensemble. That is, for any two group-generated ensembles $A$ and $B$ corresponding to groups $G_A$ and $G_B$, with $G_A \neq G_B$ but $|G_A| = |G_B|$, if their respective Gram matrices $\mathcal{G}^{G_A}$ and $\mathcal{G}^{G_B}$ share the same eigenvalues, then the exclusion probabilities for both \gls{me} and \gls{ze} protocols are identical. This group-independence property is a cornerstone of the present study.

@@

In this project, we will extend this result by demonstrating that the perfect exclusion condition Eq. (\ref{equationPerfectExclusionCondition}) provides a lower bound for the success probability of exclusion even for non-group-generated ensembles. Since this bound equals 1, it implies that any ensemble whether or not group-generated that satisfies condition Eq. (\ref{equationPerfectExclusionCondition}) can achieve perfect exclusion. Thus, the perfect exclusion zone is not exclusive to group-generated ensembles, but also includes all ensembles that fulfill the given condition.

@@

In the non-perfect exclusion regime, the analytical results for group-generated ensembles provide lower bounds on the success probabilities, and upper bounds on the error probabilities for general ensembles.

@@

Since $U(n)$ is compact, and assuming continuity of the exclusion probabilities with respect to $U$ under a given matrix norm, for each of these unitary matrices there exist arbitrarily small neighborhoods $\mathcal{B}(U_{\mathbb{Z}/4\mathbb{Z}})$ and $\mathcal{B}(U_{\mathbb{Z}/2\mathbb{Z} \times \mathbb{Z}/2\mathbb{Z}})$ such that the exclusion probabilities within these neighborhoods approximate the analytical result as closely as desired. However, the existence of multiple such matrices, and intersecting neighborhoods, can complicate numerical optimization and analysis.

In contrast, it is usefull to consider the case where the ensemble contains a prime number $p$ of states. In this scenario, the situation simplifies considerably, since there exists only one group, up to isomorphism, of order $p$, namely $\mathbb{Z}/p\mathbb{Z}$.

This result follows immediately from Lagrange's theorem. The theorem guarantees the order of any element in a finite group\footnote{The order of an element $\tau$ in a group $G$ is the smallest positive integer $m$ such that $\tau^m = e$, where $e$ denotes the identity/neutral element of the group.} must divide the order of the group. Since $p$ is a prime number, the only positive integers dividing $p$ are either $1$ or $p$ itself. Therefore, the order of any group element must be either $1$ or $p$. An element of order $1$ must be the identity, so every non-identity element in the group must have order $p$. This implies that all non-identity elements generate the entire group, i.e., the group is cyclic. As a result, any group of prime order $p$ is necessarily cyclic and isomorphic to $\mathbb{Z}/p\mathbb{Z}$.

@@

\subsection{Haar Measure}
\hspace{20pt}In this section, we introduce the concept of a measure on a topological space and define the Haar measure. This is one of the key theoretical tools underlying this project, as we use the Haar measure to ensure a uniform and unbiased sampling of ensembles when generating Gram matrices for general ensembles.

>>>>

\documentclass[12pt,letterpaper]{article}
\usepackage{fullpage}
\usepackage[top=1.5cm, bottom=3.5cm, left=2.2cm, right=2.2cm]{geometry}
\usepackage{amsmath,amsthm,amsfonts,amssymb,amscd, esint}
\usepackage{lastpage}
\usepackage{enumerate}
\usepackage{fancyhdr}
\usepackage{mathrsfs}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{hyperref}
\usepackage[english]{babel}
\usepackage{lipsum}
\usepackage[table,xcdraw]{xcolor}
\usepackage{enumitem}
\usepackage{overpic}
\usepackage{float}
\usepackage{yfonts}
\usepackage{braket}
\usepackage{dsfont}
\usepackage{tikz}
\usepackage{wrapfig}
\usepackage{url}
%\usepackage{natbib}
\usepackage[numbers,square]{natbib}
\bibliographystyle{unsrtnat}
%\usepackage[style=numeric,sorting=none]{biblatex}

\usepackage[normalem]{ulem}
\usepackage{multicol}
\useunder{\uline}{\ul}{}

@@

\thispagestyle{empty}
\mbox{} 
\newpage
\thispagestyle{empty}
\vspace*{\fill} % Push content to vertical center
\begin{flushright}
    \emph{Ab ovo usque ad mala.}\\[1em]
    \textbf{Horace}
\end{flushright}
\vspace*{\fill} 
\newpage
\thispagestyle{empty}
\mbox{} 
\newpage
\pagestyle{empty}
\headheight 35pt
\lhead{\NetIDb}    
\rhead{\subject}
\cfoot{}
\headsep 1.5em

\section*{abstract}
\vspace{1cm}
\hspace{20pt}Unlike the well-studied problem of \gls{qsd}, which aims to identify an unknown quantum state from a known ensemble, \gls{qse} focuses on confidently ruling out one or more possible candidates. This approach becomes particularly valuable in scenarios where exclusion can be achieved without error.

This project investigates \gls{qse} under two principal protocols, \gls{me}, which aims to minimize the risk of incorrectly excluding the true state, and \gls{ze}, which seeks to guarantee perfect exclusion at the cost of potentially inconclusive outcomes. Building on recent analytical results for group generated ensembles in these protocols \cite{MainPaper}, we carry out a numerical study of \gls{qse} task using \gls{sdp} techniques, analyzing general ensembles and benchmarking them against the  group generated cases, i.e. the lower bound.

We begin by introducing the mathematical formulation of the \gls{qse} problem and the theoretical tools necessary to interpret our results. We also describe a protocol for generating random generic ensembles in a homogeneous manner according to the Haar measure.

Our findings confirm that the analytical bounds established for group generated ensembles serve as effective lower bounds for more general cases. We present detailed results for three-state ensembles and examine the statistical distribution of correct exclusion probabilities for randomly sampled ensembles. Furthermore, we provide theoretical insights and intuition to support the observed behavior.

These results underscore the unique role of \gls{qse} in quantum information theory, particularly in contexts where partial state inference and hypothesis elimination are more feasible or desirable than full state identification.

\newpage
\thispagestyle{empty}
\cfoot{}
\tableofcontents

\newpage
\pagestyle{fancyplain}
\headheight 35pt
\lhead{\NetIDb}    
\rhead{\subject}
\cfoot{\small\thepage}
\headsep 1.5em
\setcounter{page}{1}
%Keywords: \emph{SDP, Quantum state exclusion , \textcolor{blue}{Add keywords}}.
\section{Introduction}	\label{sectionIntroduction}

@@

In the quantum realm, a similar line of reasoning can be applied in some tasks in this field. One of the fundamental tasks in quantum information theory is the identification of an unknown quantum state known to belong to a given ensemble of possible states. This task is referred to as Quantum State Discrimination (\gls{qsd}), and it has been extensively studied in the literature \cite{DiscriminationArticle}. However, instead of aiming to identify the exact state, we may take another approach by applying the hypothesis exclusion idea. That is, given a quantum state known to be one among a certain ensemble, the goal becomes to rule out the possibility of the unknown state being a specific member of that ensemble. This alternative task is known as Quantum State Exclusion (\gls{qse}) \cite{PhysicsExclusionSource}, and it offers a distinct framework with its own set of advantages \cite{piani2020quantum} with respect to its counterpart \gls{qsd}.

We can approach \gls{qse} task under several protocols, depending on the context within we perform the hypothesis exclusion we can prefer one or another. In this project, we focus on two main protocols, Minimum Error (\gls{me}) and Zero Error (\gls{ze})\footnote{Also known as \emph{unambiguous exclusion}.}, i.e. the protocols for which analytical results have been presented \cite{MainPaper}. In the \gls{me} scenario, the goal is to minimize the probability of mistakenly excluding the actual prepared state. In contrast, the \gls{ze} approach seeks to exclude a state with absolute certainty, although this procedure sometimes yields an inconclusive result when perfect exclusion cannot be accomplished.

While \gls{qsd} has seen significant progress over the years \cite{bae2015quantum, helstromBook}, \gls{qse} has remained relatively underexplored until recently. New analytical results have emerged \cite{MainPaper} for both the \gls{me} and \gls{ze} protocols. In particular, when the ensemble is generated through the action of a finite algebraic group, the problem becomes more tractable, highlighting \gls{qse} as a worthwhile object of study.

Although the tasks of exclusion and discrimination coincide for ensembles containing only two states\footnote{Since for the two states case excluding one necessarily implies identifying the other.}, when dealing with ensembles of three or more states, the two problems diverge in both approach and complexity. In order to further motivate the interest in \gls{qse} and discuss the main differences, we include a brief comparison of both tasks and protocols  with \gls{qsd} in Appendix \ref{appendixComparisonQSDvsQSE}. One of the key features of \gls{qse} is the possibility of achieving \emph{perfect exclusion}, i.e. confidently ruling out a candidate state with complete certainty that it is different from our given unknown state. In contrast, \emph{perfect discrimination} is generally not achievable, meaning that one cannot identify the unknown state with absolute certainty in most settings\cite{OptimalitySRM}.

@@

Building upon the recent results for the probability of a correct exclusion for group generated ensembles \cite{MainPaper},\footnote{A proper definition of a \emph{group generated ensemble} will be given in the \emph{Group generated ensembles} section \ref{sectionGroupGeneratedEnsemble}, for the time being, we can identify them as ensembles that comprehend set of states that exhibit a certain symmetry.} this project undertakes a numerical investigation into \gls{qse}, highlighting the analytical results from group generated ensembles as lower bounds for more general cases. The project comences with a preliminaries section that outlines the mathematical framework underlying the \gls{qse} task, along with the essential tools used in our analysis. We then present results for the probability of correct hypothesis exclusion in ensembles consisting of three quantum states, the simplest case. Furthermore, we explore the distribution of correct exclusion probabilities for randomly generated ensembles sampled according to Haar measure.\footnote{The importance of the generic ensembles generation being performed according to Haar measure is a key for the numerical analysis and it is discussed in the \emph{Methodology} section (\ref{sectionMethodology})}

To carry out our numerical analysis, we employ \gls{sdp} techniques,\cite{vandenberghe1996semidefinite} which enable us to formulate the \gls{qse} task as a convex optimization problem. The \gls{sdp} approach has proven particularly effective for problems of this nature \cite{bandyopadhyay2014conclusive}. Finally, we conclude by providing some theoretical intuition and insights based on our results.

@@

Unlike projective measurements, the operators $\Pi_i$ in a \gls{povm} are not necessarily orthogonal projectors. This allows the number of measurement outcomes to exceed the dimensionality of the Hilbert space, extending the previos orthogonal measurement to a more generic status and making POVMs particularly useful in tasks such as \gls{qse} and \gls{qsd}.

@@

\hspace{20pt}Let $\left\{(\rho_i, \eta_i)\right\}_{i=1}^n$ be an ensemble of $n$ quantum states, where each $\rho_i$ denotes a pure state density matrix\footnote{This formulation holds true for mixed states but the project will only consider the pure state scenario.}, i.e., $\rho_i = \ket{\psi_i}\bra{\psi_i}$, and $\eta_i$ represents the prior probability of occurrence of the state $\rho_i$. Let $\tilde{\rho}$ be a state from this ensemble. Our objective is to develop a procedure to identify another state $\rho_j \in \left\{\rho_i\right\}_{i=1}^n$, such that $\rho_j \neq \tilde{\rho}$.

Additionally, let $\left\{\Pi_i\right\}_{i=1}^n$ be a set of \glspl{povm} acting on the Hilbert space $\mathcal{H}$ of the quantum system. Here we present the two studied protocols for \gls{qse}: Minimum Error (\gls{me}) and Zero Error (\gls{ze}).

As previously discussed the goal of the \gls{me} protocol is to minimize the probability of incorrectly excluding the unknown state from our hypothesis. If we formulate the error probability as an \gls{sdp}, we obtain,\footnote{Note that SDP formulations for quantum state discrimination may differ from exclusion tasks by interchanging minimization and maximization objectives.}
\begin{align*}
	P_{\text{\gls{me}}}^e = \min_{\left\{\Pi_i\right\}} &\sum_{i=1}^n \eta_i \tr(\Pi_i \rho_i),\\
	\text{subject to} \quad & \sum_{i=1}^n \Pi_i = \mathds{1}, \quad \Pi_i \geq 0 \quad \forall i \in \{1, \dots, n\}.
\end{align*}

@@

In the case of the \gls{ze} protocol, the objective is to exclude with total certainty a state known to be different from the unkown state. However, the protocol might lead to inconclusive result. In order to do so, the \gls{povm}s must also satisfy an unambiguity condition, i.e. each measurement operator $\Pi_i$ must be orthogonal to the corresponding state $\rho_i$. In other words,
\begin{align*}
\tr(\Pi_i \rho_i) = 0 \quad \forall i \in \{1, \dots, n\}.
\end{align*}

@@

If the measurement yields the outcome "?" , the result is inconclusive.

@@

This reformulation reveals that if two ensembles $A$ and $B$ share the same Gram matrix, then their exclusion problems under both the \gls{me} and \gls{ze} protocols are equivalent. That is, their optimal success and error probabilities coincide. Consequently, we focus on the Gram matrix to study exclusion problems rather than the explicit state representations, as it encodes all the relevant information.

@@

\hspace{20pt}As we have mentioned before (Section \ref{sectionIntroduction}), analytical results have been found for the exclusion task in both protocols \gls{me} and \gls{ze} for ensembles that exhibit a certain degree of symmetry\cite{MainPaper}, that is for \emph{group generated} ensembles. In this section we present a proper definition for a group generated ensemble as well as a brief discussion on this ensembles generated by a cyclic group, a key type of group generated ensembles in our analysis.

@@

forms a group generated ensemble based on the cyclic group $\mathbb{Z}/n\mathbb{Z}$. This type of ensemble is of particular interest in our study and will be explored in more detail at the end of this section. Nonetheless, in Appendix \ref{appendixComputationGroupGeneratedGramMatrices} we present a general procedure to compute the Gram Matrix for any group generated ensemble.

\subsubsection*{The $\mathbb{Z}/n\mathbb{Z}$ group generated ensemble}

\hspace{20pt}Here we discuss the particular case for a $\mathbb{Z}/n\mathbb{Z}$ group generated ensemble a key type of ensemble for our project.\footnote{The relevance of this type of ensembles relies on the fact that for a prime number of states, the possible group generated group ensemble } Let $\mathcal{U} \in U(n)$ be an $n \times n$ unitary matrix satisfying $\mathcal{U}^n = \mathds{1}$, and let $\ket{\psi}$ be the seed state. The Gram matrix $\mathcal{G}^{\mathbb{Z}/n\mathbb{Z}}$ elements associated with the ensemble $\left\{\mathcal{U}^i\ket{\psi}\right\}_{i=0}^{n-1}$ are,
\begin{align*}
	\mathcal{G}^{\mathbb{Z}/n\mathbb{Z}}_{i,j} = \sqrt{\eta_i \eta_j}\braket{\mathcal{U}^i\psi | \mathcal{U}^j\psi} = \sqrt{\eta_i \eta_j}\braket{\psi | \mathcal{U}^{j-i} | \psi} = \sqrt{\eta_i \eta_j}\braket{\mathcal{U}^{j-i}}_{\psi},
\end{align*}

@@

For instance, the $\mathbb{Z}/4\mathbb{Z}$ group generated gram matrix with equal prior probabilities is,

\begin{align*}
\mathcal{G}^{\mathbb{Z}/n\mathbb{Z}} = \frac{1}{4}\begin{pmatrix}1 & c_1 & c_2 & c_1^* \\c_1^* & 1 & c_1 & c_2  \\c_2 & c_1^* & 1 & c_1\\c_1 & c_2 & c_1^* & 1  \\\end{pmatrix},
\end{align*}

where we have denoted $\braket{\mathcal{U}^i}_{\psi}$ as $c_i$. This shows the circulant structure of $\mathcal{G}$ for an equal prior probability case, where each row is a cyclic permutation of the previous one. In other words, the Gram matrices of $\mathbb{Z}/n\mathbb{Z}$ goup generated ensembles are circulant matrices when taking equal priors probabilities. This structure is particularly useful, as circulant matrices can be diagonalized by the discrete Fourier basis, greatly simplifying many computations~\cite{circulantMatrices}.

@@

Finally, we denote the cardinality of a finite group $G$, i.e. the number of its elements of the group, as $|G|$. In scenarios with uniform prior probabilities, this implies,
\begin{align*}
	\eta_i = \frac{1}{|G|}\quad \forall i\in\{1,...,n\}.
\end{align*}

\newpage
\subsection{Previous Results}\label{sectionPreviousResults}

\hspace{20pt}Significant progress has been made in the study of the state exclusion task for group generated ensembles. In particular, the work \emph{Quantum State Exclusion for group generated Ensembles of Pure States}\cite{MainPaper} presents exact analytical expressions for the success and error probabilities under both the Minimum Error (\gls{me}) and Zero Error (\gls{ze}) exclusion protocols. In this section we present this results and comment some key aspects of this results regarding our project, in particular we present the concept of \emph{perfect exclusion region} and we discuss the case of ensembles with a prime number of states.

Let $\mathcal{G}^G$ denote the Gram matrix of a group generated ensemble $\left\{ \left( \rho_i, \frac{1}{|G|} \right) \right\}_{i=1}^{|G|}$, where $G$ is a finite group of size $|G|$, and let $\{\lambda_i\}_{i=1}^{|G|}$ be the set of eigenvalues of $\mathcal{G}^G$, the Gram Matrix of the group. Assume the eigenvalues $\left\{\frac{\lambda_i}{|G|}|1\leq i \leq n\right\}$ are ordered increasingly,
\begin{align*}
	i \leq j \quad \Leftrightarrow \quad \lambda_i \leq \lambda_j, \quad \forall i,j \in \{1, \dots, |G|\},
\end{align*}

@@

The exclusion probabilities under the Minimum Error protocol are given by,
\begin{align*}
	P_{\gls{me}}^s &= \min\left\{ 1, 1 - \left( \frac{ \sqrt{\lambda_{|G|}} - \sum_{i<|G|} \sqrt{\lambda_i} }{|G|} \right)^2 \right\}.
\end{align*}

For the Zero Error protocol, the corresponding expressions are,
\begin{align*}
	P_{\gls{ze}}^s &= \min\left\{ 1, 1 - \frac{ \sum_{i=1}^{|G|} \sqrt{\lambda_i} \left( \sqrt{\lambda_{|G|}} - \sum_{j<|G|} \sqrt{\lambda_j} \right) }{|G|} \right\}.
\end{align*}

A key feature of these results is their independence from the specific group generating the ensemble. That is, for any two group generated ensembles $A$ and $B$ corresponding to groups $G_A$ and $G_B$, with $G_A \neq G_B$ but $|G_A| = |G_B|$, if their respective Gram matrices $\mathcal{G}^{G_A}$ and $\mathcal{G}^{G_B}$ share the same eigenvalues, then the exclusion probabilities for both \gls{me} and \gls{ze} protocols are identical. This group-independence property is a cornerstone of the present study.

@@

In this project, we will extend this result by showcasing that the perfect exclusion condition Eq. (\ref{equationPerfectExclusionCondition}) provides a lower bound for the success probability of exclusion even for non-group generated ensembles. Since this bound equals 1, it implies that any ensemble whether or not group generated that satisfies condition Eq. (\ref{equationPerfectExclusionCondition}) can achieve perfect exclusion. Thus, the perfect exclusion zone is not exclusive to group generated ensembles, but also includes all ensembles that fulfil the given condition.

@@

In the non-perfect exclusion regime, the analytical results for group generated ensembles provide lower bounds on the success probabilities, and upper bounds on the error probabilities for general ensembles.

@@

Since $U(n)$ is compact, and assuming continuity of the exclusion probabilities with respect to $U$ under a given matrix norm, for each of these unitary matrices there exist arbitrarily small neighbourhoods $\mathcal{B}(U_{\mathbb{Z}/4\mathbb{Z}})$ and $\mathcal{B}(U_{\mathbb{Z}/2\mathbb{Z} \times \mathbb{Z}/2\mathbb{Z}})$ such that the exclusion probabilities within these neighborhoods approximate the analytical result as closely as desired. However, the existence of multiple such matrices, and intersecting neighbourhoods, can complicate numerical optimization and analysis.

In contrast, it is useful to consider the case where the ensemble contains a prime number $p$ of states. In this scenario, the situation simplifies considerably, since there exists only one group, up to isomorphism, of order $p$, namely $\mathbb{Z}/p\mathbb{Z}$.

This result follows immediately from Lagrange's theorem\cite{LagrangeTheoreme}. The theorem guarantees the order of any element in a finite group\footnote{The order of an element $\tau$ in a group $G$ is the smallest positive integer $m$ such that $\tau^m = e$, where $e$ denotes the identity/neutral element of the group.} must divide the order of the group. Since $p$ is a prime number, the only positive integers dividing $p$ are either $1$ or $p$ itself. Therefore, the order of any group element must be either $1$ or $p$. An element of order $1$ must be the identity, so every non-identity element in the group must have order $p$. This implies that all non-identity elements generate the entire group, i.e., the group is cyclic. As a result, any group of prime order $p$ is necessarily cyclic and isomorphic to $\mathbb{Z}/p\mathbb{Z}$.

@@

\subsection{Haar Measure}
\hspace{20pt}In this section, we introduce the concept of a measure on a topological space and define the Haar measure. This is one of the key theoretical tools underlying this project, as we use the Haar measure to ensure a uniform sampling of ensembles when generating Gram matrices for general ensembles.
